Measuring the Influence of Trollbots in Twitter during the 2018 US Midterm Election

### 2019-12-13 paper.tex
Updating the paper

\documentclass[letterpaper]{article}
\usepackage{aaai19}
\usepackage{times} 
\usepackage{helvet}
\usepackage{courier} 
\usepackage[hyphens]{url}
\usepackage{graphicx}
\urlstyle{rm}
\def\UrlFont{\rm}
\usepackage{graphicx}
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
% custom packages%
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{subcaption}

\pdfinfo{
/Title (Measuring the Influence of Twitter Bots during the 2018 US Midterm Election)
/Author ()
}

@@

\title{Measuring the Influence of Twitter Bots during the 2018 US Midterm Election}

@@

\begin{abstract}
    Use of bots in social media and its consequences has been a subject of great public interest. In this study, we measure the influence of bots during 
    the 2018 US Midterm Elections. Between October 31, 2018 and November 7, 2018 we collected 6.48M tweets and 1.6M profile data from Twitter. We used 
    Machine Learning techniques to separate bot and human accounts and found that 1\% of the
    top bots influenced the creation of 11\% of all the tweets. Additionally, for URLs which were retweeted more than ten times in our dataset, we found that the 
    total influence percentage of bots was double their size. 
    Influence is a number we give to a tweet which measures 
    the expected number of time the tweet gets retweeted over all possible diffusion scenarios. Influence is converted to percentage by using the sum of user wise influence 
    throughout the dataset.
\end{abstract}

@@

Twitter is a social media where short messages get exchanged. Messages shared by a user are shown to
their followers and to followers of those who share the message. These messages are called Tweets. Sharing can be done by retweeting
(sharing) or replying to the tweet. When Alice replies/retweets, a tweet made by Bob, Alice and Bob's followers will see the tweet. In this
way, messages spread. Having massive followers, getting many retweets, getting retweeted by people with massive followers or advertising can make a tweet viral.\par

People use social media, at least in part, to form an opinion about lifestyle, health, politics, and purchase \cite{varol2017early}. 
Due to the power of social media in influencing opinion, various ethical and unethical techniques have been devised to reach a big audience. 
One such unethical technique is the use of bots. Social bots are account controlled wholly or partially by a computer algorithm. 
These bots can generate content and interact with human users often imitating humans \cite{ferrara2016rise}.
A US special counsel investigation found that Twitter accounts of US personas were being operated as a botnet (a network of bots) to amplify content. 
During the 2016 US Presidential Election, they chose candidates to support and oppose \cite{luceri2019red,mueller_investigation}. 
These propaganda bots were found to be high volume, multichannel, rapid, continuous, repetitive, inaccurate, and inconsistent \cite{paul2016russian}. \par

Past studies have explored the political polarization \cite{munger2017don,rizoiu2018debatenight,gruzd2014investigating,bovet2019influence}, origin of bots 
\cite{zannettou2019let,zannettou2019characterizing} and the spread of fake news \cite{vosoughi2018spread,shao2018spread,lazer2018science,bovet2019influence,grinberg2019fake}
on twitter. Previous US Election has been studied before. Most studies focus on 2016 Election \cite{bovet2019influence,rizoiu2018debatenight,bessi2016social,howard2018algorithms,howard2016bots}. \cite{deb2019perils} 
studied the 2018 Midterm Election.\par

Few studies perform a large scale quantitative study of the role played by bots during a significant event. To fill the gaps in the literature, 
we measured the influence of bots during the 2018 US Midterm Election. \par

@@

\begin{itemize}
    \item 10.5\% of the users in the datasets were bots. They made 12.6\% of the tweets and created 22\% of the total influence.
    \item On Average, bots were twice as influential as humans
    \item Top 1\% of the bots influenced the creation of 11\% of all the retweets during this time.
    \item In average a single URL was shared 27.84 times in 1.97 cascades. URL cascade is formed when a Tweet mentions a URL. Any further retweet belong to the same cascade.
    \item For URLs which were shared more than ten times, the influence percentage of bots was double their size.
    \item 81.64\% of all interaction (retweets and replies) took place between humans, and 2.87\% took place between bots. 6.37\% was bot-human, and 9.14\% was human-bot interaction.
    \item Most of the codes and notebooks are shared publicly in https://github.com/warproxxx/2018Midterm to help future researchers and ensures transparency.
\end{itemize}

*
\section{Related Work}
\label{sec:related}

\subsection{Social Media and Bots}
People are spending more and more time on social media. Globally, people spend 2 hours and 23 minutes on average in Social Media. 40\%  people use social media to stay up to date 
with news and current events \cite{2019_social_flagship_report}. 68\% of American adults get their news from
social media while 42\% find news on social media to be mostly accurate \cite{pew_research_news}.\par

Twitter is not the most popular social media. 12\% of US adults get news from Twitter compared to Facebook's 43\% \cite{pew_research_news}. However, 
Twitter is the number one platform for government leaders. 97\% of all UN members have an official presence in Twitter \cite{twiplomacy}. Furthermore, Twitter API provides easy access to data.
Consequently, Twitter has been widely studied \cite{rizoiu2018debatenight,grinberg2019fake,bovet2019influence,morstatter2018alt,munger2017don,gruzd2014investigating,zannettou2019characterizing,howard2016bots}.
\par

@@

Different types of bots exist in social media. A simple bot will post predetermined messages at predetermined intervals \cite{haustein2016tweets}. Bots are also used to increase followers
and make an account appear popular \cite{cresci2015fame}. Botnets are a network of bots working together. The word Botnet is also used to refers to a network of compromised computer. However, in this report, 
the word botnet refers to a network of social media bots working together to make an influence and should not be confused with the other definition. Many botnets use a hybrid human/automation 
approach \cite{grimme2018changing}. Botnets have been used to promote spam \cite{ferrara2018measuring}, which seem to shifted to social media due to the effectiveness of spam filter \cite{gao2010detecting,chu2012detecting,ferrara2018measuring}, 
manipulate stocks \cite{ferrara2015manipulation}, manipulate elections \cite{morstatter2018alt}, 
and for various other purposes \cite{abokhodair2015dissecting}.
Botnets were detected as early as 2010 US Midterm Election \cite{mitter2014categorization}. Over time, their use and study have increased. 
Studies have detected and analyzed bots in the 2017 German Federal Election \cite{morstatter2018alt} ,2017 French Presidential Election \cite{ferrara2017disinformation} and during 
various US Elections \cite{mitter2014categorization,bovet2019influence,rizoiu2018debatenight,bessi2016social,howard2018algorithms,howard2016bots,deb2019perils}. In Elections, 
they have been used to support candidates \cite{luceri2019red}, attack people \cite{mueller_investigation} and spread fake news \cite{vosoughi2018spread,grinberg2019fake}. 
Political bots are active beyond the election time. \cite{stewart2018examining} found that Russian bots infiltrated both right and left-leaning communities and spread different narratives 
\cite{mueller_investigation}. \par

In Twitter, bots share content 
in multiple channel \cite{paul2016russian}. This activity is in line with literature which shows that information from multiple sources appears more trustworthy than a single one \cite{harkins1981multiple}. 
 Bots attack people. Attacking trustworthiness has been shown to diminish the credibility of the Original Poster \cite{pornpitakpan2004persuasiveness}. 
Bots are continuous, repetitive, but most of the times, they are also far from reality \cite{paul2016russian}. The high volume seems to ensure early exposure and \cite{petty1994think} shows that early
 first impression is more likely to be accepted by the brain. Due to a phenomenon called Sleeper Effect, where information gets disassociated from the source while remembering \cite{underwood1998memory,paul2016russian}
 low credibility sources can have persuasive power to unbiased or neutral people. As a boost to sleeper effect, people are 31\% more likely to remember what they see on twitter, 
 compared to the normal web \cite{twitter_remember}. \par

\subsection{Bot Detection}
Most studies \cite{rizoiu2018debatenight,yang2019arming,shao2018spread} use Botometer\cite{davis2016botornot,yang2019arming} for bot detection.
Botometer's Machine Learning algorithm provides an account-level bot classification. Botometer classifies account partially and completely automatized as bots. 
Botometer has some false positives. Its detection of organizational accounts as bots have been highly criticized \cite{varol2017early} \cite{botometer_tweet}. However, even with its false positive, 
Botometer has no transparent competitor and 
remains the most accurate bot detection system in academia. \par

Botometer uses more than a thousand features created from temporal activity, network structure, content analysis, sentiment analysis, and user profile data to
determine a bot score \cite{davis2016botornot,yang2019arming}. \cite{bessi2016social} illustrated that profile customization, geographical metadata, and activity statistics provided the strongest signals for bot detection. \cite{ferrara2017disinformation} and \cite{kudugunta2018deep} used various machine learning techniques on these features. \cite{ferrara2017disinformation}
obtained 93\% accuracy with an AUC-ROC score of 92\% in their best model using Random Forest classifier. \cite{kudugunta2018deep} used 3,000 labeled examples to train a system with an AUC greater than 99\%
using Adaboost Classifier with Over Sampling and Undersampling of data using the SMOTENN algorithm. \cite{kudugunta2018deep} also presented a tweet level classification using contextual LSTM
classifier with tweet metadata and deep learning. \par

Most studies show that 10-20\% in social media are bots. In earlier studies, the number was on the higher side. 18\% of users in \cite{bovet2019influence}'s study were bots.
A more recent study by \cite{deb2019perils} states that their number may have dropped.

@@

\cite{shao2018spread} also found that articles spread 
mostly through tweets and retweets and much lesser from replies. Their study showed that people do not discriminate between resources shared by humans and bots. 
\cite{shao2018spread} and \cite{grinberg2019fake} found super spreaders. Moreover,  \cite{grinberg2019fake} found that  1\% of individuals accounted for 80\% of fake news exposure. 
\cite{varol2017early} also found that 2\% user accounts were responsible for 60\% of the conversation. \par

\cite{rizoiu2018debatenight} found that social bots were 2.5 times more influential than humans. They introduce a scalable algorithm for estimating user influence in retweet cascades. 
For each tweet in an information cascade, it uses time of post and the number of followers to determine the probability of whether a tweet is a retweet of another. They then find influence which can be used 
to find out how many users the tweet possibly influenced to retweet. It was tested successfully in artificial social media. Other studies have attempted to measure influence before this. 
\cite{weng2010twitterrank} used eigenvector centrality of the connection to measure influence, but it is not scalable into big cascades. There are other methods like 
\cite{rodriguez2011uncovering,cho2013latent,linderman2014discovering} but they either have scaling issue or require full diffusion graph, which Twitter does not provide.

\section{Methods}
\label{sec:method}
\subsection{Data Collection}
Twenty-nine manually identified keywords, and names of all 53 House candidates and 27 Senate candidates were used to collect live data from Twitter. The manually defined keywords are in 
Table \ref{tab:manual-keywords}

\begin{table}[]
    \centering
    \begin{tabular}{ll}
    • 2018Midterms & • resist \\
    • 2018MidtermElections & • VOTE \\
    • Election2018 & • GAEarlyVoting \\
    • ElectionDay & • EarlyVoting \\
    • MAGA2018 & • plus1 \\
    • MAGA & • IVoted \\
    • Trump2020 & • WinBlue \\
    • AmericaFirst & • WinRed \\
    • TheResistance & • BlueWave \\
    • WalkAwayFromRepublicans2018 & • RedWave \\
    • VoteThemOut & • republican2018 \\
    • 2018Senate & • democrat2018 \\
    • VoteRed & • Republican \\
    • Voteblue & • Democrat \\
    • WalkAwayFromDemocrats2018 & 
    \end{tabular}
    \caption{Manual Keywords}
    \label{tab:manual-keywords}
\end{table}

Twitter provides Streaming API, Search API, and a premium Firehose API to provide access to tweets. Streaming API provides limited access to live data (thus also called 1\% API)
 while Firehose API
provides complete data. However, due to the associated costs, Firehose API was not an option in this study. \cite{morstatter2013sample} compares the Streaming API with Firehose API.
They found that Streaming API was nearly as good as a random sample of Firehose API when the dataset was large enough. Although the 1\% API was not as good as a 1\% random sample from the Firehose API in 
all of their tests, it estimated the top hashtags correctly when the data was large enough. It recommended the creation of specific parameters and using a large sample. 
Due to the possible issues in Streaming API, \cite{bessi2016social} recommends using Twitter Search API.

\par In this study, we use Streaming API with many keywords as suggested by \cite{morstatter2013sample} to collect a large dataset. While collecting data, information about the tweeter and the tweet was collected.

@@

We collected Tweets from October 31, 2018 (1 AM UTC) to November 7, 2018 (midnight UTC). On July 18, 2019, we used Twitter API to rescrape the top 100k most retweeted which did not 
exist in our dataset.

\subsection{Data Processing}
First, we manually removed irrelevant tweets. We created keywords find and remove these irrelevant data.\par

We used SentiStrength and VADER Sentiment analysis in the Tweets. SentiStrength \cite{thelwall2010sentiment} is used to annotate short, informal tweet like texts \cite{bessi2016social}. 
It can capture positive and negative emotions at an accuracy of 60.6\% and 72.8\%. SentiStrength gives a positive and negative score between 0 and 4. The positive sentiment is subtracted 
from the negative like in \cite{bessi2016social} to get a whole number. VADER is a rule-based sentiment classification designed for use in social media data. VADER was found to outperform 
individual human raters with an F1 accuracy of 0.96 compared to 0.84 for humans \cite{hutto2014vader}.

\subsection{Bot Detection}
Detecting bots on the wild is more complicated than against a validation set \cite{bovet2019influence,varol2017early,ferrara2016rise}. The collected dataset had 1.7M 
users. Although Botometer is publicly available, getting information for 1.7M data would be expensive. So the following steps were used to train a bot detection system:

\begin{itemize}
    \item \cite{cresci2017paradigm} provided a labeled account level bot or not data with relevant features that included the ones we collected.
    We used these features to train a Random Forest Classifier and detect bots. 
    \cite{cresci2017paradigm}'s data is old, and bots evolve quickly. So further processing was required.
    \item We used the trained model on our dataset, which contains the same features, to select 40k highly probable bots.
    \item We selected 40k highly probable bots and 32k random accounts from our dataset and performed bot detection through Botometer. Botometer returns CAP score and bot score. CAP score denotes the probability of being a bot. 
    Previous studies have used bot score of 0.5 to determine bots. CAP score of 0.3 is equivalent to the score 0.5 used by researchers in these earlier studies. \cite{deb2019perils} used a threshold of 0.3 after the 
    introduction of new changes as mentioned in \cite{yang2019arming}.
    \item Grid search was performed with thresholds of 0.1, 0.2, and 0.3 for humans, 0.3, 0.5, 0.7, and 0.8 for bots and Random Forest, Neural Network
     and SMOTENN as possible ML algorithms to determine the best parameters in the training and test set
    \item From Grid search, we determined that 0.5 was the optimal bot threshold, 0.3 the optimal human threshold and Adaboost with SMOTENN preprocessing the optimal ML algorithm. Using this threshold, we selected 513 bots whose CAP score was higher than 0.5  and 
    27k human accounts whose CAP score was smaller than 0.3 for training and testing our data. We created a training-test split at a ratio of 0.9. We used the remaining 44k values that lied between 0.5 and 0.3 as a test set,
    which we name "Broader Test set" to see the models performance in thresholds that were not selected. 
\end{itemize}

We obtained an accuracy of 85\% with AUC of 0.85 in the training set, 87\% accuracy with 
an AUC of 0.8 in the test set and accuracy of 85\% with an AUC of 0.8 in the broader test set when compared with Botometers prediction. This model detected 10.5\% of the accounts as 
bots. The accuracy was lesser than in \cite{kudugunta2018deep}. It may be due to the evolution of bots, or because we compared with Botometer instead of manually labeling the training set like in  \cite{kudugunta2018deep}.

@@

 Influence percentage mean. This way, it measures the role of a user in creating new tweets. More the influence percentage, more tweets in the dataset were created by that user/user group. In 
a cascade, the original tweet is the parent tweet.

 \section{Data Analysis}
\label{sec:analysis}
\subsection{Data Captured}
After cascading and re-scraping the missing tweets, we compared the actual size of the cascades with the size we had captured using the Twitter API. This comparison has issues. 
There is a selection bias in this data as we selected cascades which were longer than 2 Tweets. Additionally, we re-scraped metrics (likes, replies and retweets) from twitter in July 2019, nine months after
 the collection. By this time, addition and removal of tweets will take place. However, we still include this analysis to enable future comparison with other researchers. \par

In our 329606 cascades, we had re-scraped the details of 154603 parent
tweets. We compared the total retweet and replies we had (when re-scraped) with the total retweets and replies on Twitter. After we added the total tweet and retweet amount in Twitter  
and compared it with the amount we captured, we found that we had captured 9.64\% of data in each cascade. \par

The biases explained
above, and the use of many keywords may be possible reasons for this high capture.

\subsection{Exploratory Data Analysis}
After the removal of irrelevant data, 6.5M Tweets by 1.7M users remained. Bots made 12.6\% of the Tweets. We removed stopwords to compare the words used by bots and humans. 
We selected words which appear more than 0.03\% times in the dataset (which was 45759 times for humans and 6538 for bots). The dominant words are in 
\ref{tab:bot-dominant-words} and  \ref{tab:humans-dominant-words}. From Table \ref{tab:bot-dominant-words}, we can see that bots dominate humans in the use of conservative words. 
Table \ref{tab:humans-dominant-words} shows that humans dominate bots in the use of slang. \par

\begin{table}
    \centering
    \begin{tabular}{|l|l|}
    \hline
    \textbf{Word} & \textbf{Percentage Difference} \\ \hline
    voteredtosaveamerica & 169.715447 \\ \hline
    patriots & 161.121157 \\ \hline
    votered & 130.514988 \\ \hline
    walkaway & 107.832423 \\ \hline
    follow & 106.300268 \\ \hline
    red & 99.139168 \\ \hline
    maga & 95.622477 \\ \hline
    redwave & 95.294118 \\ \hline
    qanon & 86.301370 \\ \hline
    dems & 65.538736 \\ \hline
    usa & 60.909091 \\ \hline
    via & 57.861635 \\ \hline
    god & 56.432247 \\ \hline
    economy & 51.132075 \\ \hline
    \end{tabular}
    \caption{Words used more often by bots}
    \label{tab:bot-dominant-words}
\end{table}

\begin{table}
    \centering
    \begin{tabular}{|l|l|}
    \hline
    \textbf{Word} & \textbf{Percentage Difference} \\ \hline
    bunch & 58.440047 \\ \hline
    yall & 52.095130 \\ \hline
    bc & 45.844156 \\ \hline
    shot & 45.454545 \\ \hline
    woman & 43.218954 \\ \hline
    children & 41.602787 \\ \hline
    line & 41.316348 \\ \hline
    school & 41.313559 \\ \hline
    cruz & 39.701074 \\ \hline
    points & 39.238095 \\ \hline
    men & 39.154161 \\ \hline
    put & 35.545906 \\ \hline
    stay & 35.414725 \\ \hline
    ted & 35.111111 \\ \hline
    \end{tabular}
    \caption{Words used more often by humans}
    \label{tab:humans-dominant-words}
\end{table}

We did not find any significant difference between the sentiment of humans and bots. Then we analyzed the 329606 cascades in the dataset to calculate human-bot interaction. Out of all interactions, we made the following observations:

\begin{itemize}
    \item Tweets made by bots was retweeted by bots 2.87\% or 120k times.
    \item Tweets made by bots were retweeted by humans 6.34\% or 265k times. 
    \item Tweets made by humans were retweeted by bots 9.14\% or 382k times. 
    \item Tweets made by humans were retweeted by humans 81.64\% or 3.4M times. 
\end{itemize}

\begin{figure}
    \centering
    \begin{subfigure}[b]{1\linewidth}
      \includegraphics[width=\linewidth]{images/bots_furtherretweeted.png}
      \caption{Bots Parents}
      \label{fig:further_retweet_bots}
    \end{subfigure}
    \begin{subfigure}[b]{1\linewidth}
      \includegraphics[width=\linewidth]{images/humans_furtherretweeted.png}
      \caption{Humans Parents}
      \label{fig:further_retweet_humans}
    \end{subfigure}
    \caption{Further Retweet}
    \label{fig:further_retweet}
\end{figure}

>>>>

\documentclass[letterpaper]{article}
\usepackage{aaai19}
\usepackage{times} 
\usepackage{helvet}
\usepackage{courier} 
\usepackage[hyphens]{url}
\usepackage{graphicx}
\usepackage{float}
\urlstyle{rm}
\def\UrlFont{\rm}
\usepackage{graphicx}
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
% custom packages%
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{subcaption}

\pdfinfo{
/Title (Measuring the Influence of Trollbots during the 2018 US Midterm Election)
/Author ()
}

@@

\title{Measuring the Influence of Trollbots in Twitter during the 2018 US Midterm Election}

@@

\begin{abstract}
    The use of troll bots in social media and its consequences has been a subject of great public interest. In this study, 
    we measure the influence of trollbots during 
    the 2018 US Midterm Elections. We selected 2.94M tweets and 870k 
    profile data posted between October 31, 2018, and November 7 using nine keywords. On December 10, 2019, 
    we rescraped Twitter to find suspended and removed accounts. We identified the suspended accounts 
    as Trollbots and used Machine Learning on the removed 
    accounts to find other trollbots. We found that the top 1\% of the Trollbots, influenced the creation of 7\% of all the tweets.
    In Twitter, Tweet or a retweet made by another user can be retweeted.
    We estimate the number of times the tweet or the retweet 
    gets retweeted by other users to calculate the influence of a Tweet.
    We added the influence of all Tweets made by a user to calculate user wise influence.\end{abstract}

@@

Twitter is a social networking service where users interact with each other using short messages called Tweet. 
Tweets spread through retweets or replies. For example, when Alice replies/retweets, a tweet made 
by Bob, Alice and Bob's followers see the tweet. \par

People use social media, at least in part, to form an opinion about lifestyle, health, politics, and purchase 
\cite{varol2017early}. To reach many audience, various ethical and unethical 
techniques have been devised. 
One such unethical technique is the use of trollbots. Here we defined Trollbots as the account which violates
Twitter's terms of service. The most common reason for the suspension is spam and abusive behaviour 
\cite{twitter}. \par

In this study, we rely on Twitter for most of the bot detection to focus on measuring the influence.

@@

\begin{itemize}
    \item 13.6\% of the users in the datasets no longer existed. 4.1\% had been suspended. Of the remaining 9.5\%, 22\% were bots. Overall, 6.19\% of the users were bots. They made 
    12\% of all the Tweets.
    \item On Average, trollbots were 53\% more influential than humans
    \item Top 1\% of the trollbots influenced the creation of 7\% of all the retweets/replies.
    \item On average, a single URL was shared 26.48 times in 1.98 cascades. When a Tweet 
    mentions a URL. Any further retweet belongs to the same cascade.
    \item For URLs that were shared more than ten times, the influence percentage of bots was 70\% greater than their 
    size.
    \item 74.18\% of all interaction (retweets and replies) took place between humans, and 1.78\% took place between 
    bots. 12.04\% was bot-human, and 12\% was human-bot interaction.
\end{itemize}

\section{Related Work}
\label{sec:related}

\subsection{Social Media, Trolls and Bots}
People are spending more and more time on social media. Globally, people spend 2 hours and 23 minutes on average in Social Media. 40\%  people use social media to stay up to date 
with news and current events \cite{2019_social_flagship_report}. 68\% of American adults get their news from
social media while 42\% find news on social media to be mostly accurate \cite{pew_research_news}.\par

12\% of US adults get news from Twitter\cite{pew_research_news}. Twitter is the number one platform for government 
leaders. 97\% of all UN members have an official presence in Twitter \cite{twiplomacy}. Furthermore, Twitter API 
provides easy access to data. Consequently, Twitter has been widely studied \cite{rizoiu2018debatenight,grinberg2019fake,bovet2019influence,morstatter2018alt,munger2017don,gruzd2014investigating,zannettou2019characterizing,howard2016bots}.
\par

@@

Different types of bots exist in social media. A simple bot will post predetermined messages at predetermined intervals \cite{haustein2016tweets}. Bots are also used to increase followers
and make an account appear popular \cite{cresci2015fame}. Here, Botnet refers to a network of social media bots working together to make an influence. Many botnets use 
a hybrid human/automation approach \cite{grimme2018changing}. Botnets have been used to promote spam \cite{ferrara2018measuring}, which seem to shifted to social media due to the effectiveness of spam filter \cite{gao2010detecting,chu2012detecting,ferrara2018measuring}, 
manipulate stocks \cite{ferrara2015manipulation}, manipulate elections \cite{morstatter2018alt}, 
and for various other purposes \cite{abokhodair2015dissecting}. Many of the suspended users are a part of Botnets.
Botnets were detected as early as 2010 US Midterm Election \cite{mitter2014categorization}. Over time, their use and
 study have increased. Studies have detected and analyzed bots in the 2017 German Federal Election \cite{morstatter2018alt} ,2017 French Presidential Election \cite{ferrara2017disinformation} and during 
various US Elections \cite{mitter2014categorization,bovet2019influence,rizoiu2018debatenight,bessi2016social,howard2018algorithms,howard2016bots,deb2019perils}. In Elections, 
they have been used to support candidates \cite{luceri2019red}, attack people \cite{mueller_investigation} and spread fake news \cite{vosoughi2018spread,grinberg2019fake}. 
Political bots are active beyond the election time. \cite{stewart2018examining} found that Russian bots infiltrated both right and left-leaning communities and spread different narratives 
\cite{mueller_investigation}. \par

In Twitter, the trollbots share content 
in multiple channels \cite{paul2016russian}. This activity is in line with literature, which shows that information 
from multiple sources appears more trustworthy than a single one \cite{harkins1981multiple}. 
 Trollbots attack people. Attacking trustworthiness has been shown to diminish the credibility of the Original Poster \cite{pornpitakpan2004persuasiveness}. 
Bots are continuous, repetitive, but most of the time, they are also far from reality \cite{paul2016russian}. 
The high volume seems to ensure early exposure and \cite{petty1994think} shows that early
 first impression is more likely to be accepted by the brain. Due to a phenomenon called Sleeper Effect, where information gets disassociated from the source while remembering \cite{underwood1998memory,paul2016russian}
 low credibility sources can have the persuasive power to unbiased or neutral people. As a boost to sleeper effect, people are 31\% more likely to remember what they see on twitter, 
 compared to the normal web \cite{twitter_remember}. \par

\subsection{Trollbot Detection}
Most studies \cite{rizoiu2018debatenight,yang2019arming,shao2018spread} use Botometer\cite{davis2016botornot,yang2019arming} for bot detection.
Botometer's Machine Learning algorithm provides an account-level bot classification. Botometer classifies account 
partially and completely automatized as bots. 
Botometer uses more than a thousand features created from temporal activity, network structure, content analysis, sentiment analysis, and user profile data to
determine a bot score \cite{davis2016botornot,yang2019arming}. \cite{bessi2016social} illustrated that profile 
customization, geographical metadata, and activity statistics provided the strongest signals for bot detection.
However, Botometer has high false positive and detects organizational accounts as bots \cite{varol2017early} \cite{botometer_tweet}.
\par

\cite{ferrara2017disinformation} and \cite{kudugunta2018deep} had success detecting bots using fewer features. 
\cite{ferrara2017disinformation} obtained 93\% accuracy with an AUC-ROC score of 92\% in their best model 
using Random Forest classifier. \cite{kudugunta2018deep} used 3,000 labelled examples to train a system with an 
AUC greater than 99\% using Adaboost Classifier with Over Sampling and Undersampling of data using the SMOTENN 
algorithm. \par

@@

\cite{shao2018spread} found that articles spread 
mostly through tweets and retweets. Their study showed that people do not discriminate between resources 
shared by humans and bots. \cite{shao2018spread} and \cite{grinberg2019fake} found super spreaders. 
Moreover, \cite{grinberg2019fake} found that 1\% of individuals accounted for 80\% of fake news exposure. 
\cite{varol2017early} also found that 2\% user accounts were responsible for 60\% of the conversation. \par

\cite{rizoiu2018debatenight} found that social bots were 2.5 times more influential than humans. They used 
Botometer to label the bots and introduced a 
scalable algorithm for estimating user influence in retweet cascades. 
For each tweet in an information cascade, it uses time of post and the number of followers to determine the 
probability of whether a tweet is a retweet of another. They then find influence which can be used 
to find out how many users the tweet possibly influenced to retweet. It was tested successfully in 
artificial social media. Other studies have attempted to measure influence before this. 
\cite{weng2010twitterrank} used eigenvector centrality of the connection to measure influence, but it is 
not scalable into big cascades. There are other methods like 
\cite{rodriguez2011uncovering,cho2013latent,linderman2014discovering}, but they either have scaling issues or 
require a full diffusion graph, which Twitter does not provide.

\section{Methods}
\label{sec:method}
\subsection{Data Collection}
Tweets containing nine manually identified keywords were selected from collected live data from Twitter. The 
keywords are:

\begin{itemize}
    \item 2018Midterms
    \item Election
    \item Election2018
    \item 2018Senate
    \item WinBlue
    \item WinRed
    \item BlueWave
    \item RedWave
    \item Republican
    \item Democrat
\end{itemize}

Twitter provides Streaming API, Search API, and a premium Firehose API to give access to tweets. 
Streaming API provides limited access to live data, while Firehose API provides complete data. However, due to the 
associated costs, Firehose API was not an option in this study. \cite{morstatter2013sample} compares the 
Streaming API with Firehose API. They found that Streaming API was nearly as good as a random sample of Firehose API 
when the dataset was large enough. Although the 1\% API was not as good as a 1\% random sample from the Firehose API 
in all of their tests, it estimated the top hashtags correctly when the data was large enough. \par

In this study, we use Streaming API. While collecting data, information about the tweeter and the tweet was collected.

@@

We collected Tweets from October 31, 2018 (1 AM UTC) to November 7, 2018 (midnight UTC). On July 18, 2019, we used 
Twitter API to rescrape the top 100k most retweeted Tweets which did not exist in our dataset.

\subsection{Data Processing}
We used SentiStrength in the Tweets to compare the sentiment. SentiStrength \cite{thelwall2010sentiment} is used 
to annotate short, informal tweet like texts \cite{bessi2016social}. 
It can capture positive and negative emotions at an accuracy of 60.6\% and 72.8\%. SentiStrength gives a positive 
and negative score between 0 and 4. The positive sentiment is subtracted from the negative like in \cite{bessi2016social} to get a whole number. SentiStrength is used instead of another 
popular solution VADER because it performed better in manual analysis.

\subsection{Trollbot Detection}
A year after the data collection, we checked twitter to find accounts that no longer existed. 117650 of the 863539
accounts did not. Of them, the 32323 suspended accounts were verified to be troll bots. The remaining 85327 
could or could not be bots as they also include accounts deactivated by users. So, we used Machine Learning in this 
data. Initially, we had used Machine Learning on the complete dataset, but we received bad accuracy using that 
mechanism. \par

We selected 95\% of suspended users and an equal amount of random non suspended users, which we conveniently refer to as humans to train a SMOTENN network using 
the profile features detailed above. It had an AUC of 0.95 in the training set and 0.93 in the test set.

@@

 \section{Data Analysis}
\label{sec:analysis}

\subsection{Exploratory Data Analysis}
Our dataset consisted of 2.94M Tweets by 870k users. After Machine Learning, 6.19\% of these accounts 
were found to be bots. They made 15.6\% of the Tweets. The mean 
sentiment of bots was -0.19, slightly more negative than the average non-bot at -0.16. We 
removed stopwords to compare the words used by bots and humans (used for convenience to refer to users who are not Trollbots). 
We selected words that appear more than 0.03\% times in the dataset (which was 41386 times for humans and 6148 for 
bots). The dominant words are in 
\ref{tab:bot-dominant-words} and \ref{tab:humans-dominant-words}. From Table \ref{tab:bot-dominant-words}, 
we can see that most of the trollbots seem to be alt-right or republican. Qanon is an alt-right conspiracy theory 
while kag refers to "Keep America Great". Other bot dominant words are also popular among the republicans.
Among humans, one incident seems to have stood out. A Florida State University (FSU) student had been 
arrested for throwing chocolate milk and the word related to it dominate the human side.

\begin{table}[h]
    \begin{tabular}{|l|l|}
    \hline
    \textbf{Word} & \textbf{\begin{tabular}[c]{@{}l@{}}Percentage\\ Difference\end{tabular}} \\ \hline
    chocolate     & 78.885052                                                                \\ \hline
    milk          & 76.540881                                                                \\ \hline
    fsu           & 76.08353                                                                 \\ \hline
    throwing      & 75.873016                                                                \\ \hline
    woman         & 59.52                                                                    \\ \hline
    arrested      & 58.247903                                                                \\ \hline
    men           & 56.640345                                                                \\ \hline
    jewish        & 55.259654                                                                \\ \hline
    nazis         & 49.34877                                                                 \\ \hline
    name          & 46.192053                                                                \\ \hline
    put           & 44.897959                                                                \\ \hline
    away          & 43.264503                                                                \\ \hline
    something     & 41.046607                                                                \\ \hline
    angry         & 40.981013                                                                \\ \hline
    think         & 40.675676                                                                \\ \hline
    \end{tabular}
    \caption{Words used more often by humans}
    \label{tab:humans-dominant-words}
\end{table}

\begin{table}[h]
    \begin{tabular}{|l|l|}
    \hline
    \textbf{Word}        & \textbf{percentage\_difference} \\ \hline
    qanon                & 32.315522                       \\ \hline
    kag                  & 31.95122                        \\ \hline
    walkaway             & 28.093948                       \\ \hline
    usa                  & 27.511962                       \\ \hline
    patriots             & 26.732673                       \\ \hline
    racism               & 26.315789                       \\ \hline
    lie                  & 25.85034                        \\ \hline
    americafirst         & 25.561798                       \\ \hline
    wwg1wga              & 25.519288                       \\ \hline
    voteredtosaveamerica & 25.504152                       \\ \hline
    low                  & 24.262295                       \\ \hline
    truth                & 24.08377                        \\ \hline
    borders              & 23.773006                       \\ \hline
    destroy              & 23.054755                       \\ \hline
    lying                & 22.685185                       \\ \hline
\end{tabular}
\caption{Words used more often by bots}
\label{tab:bot-dominant-words}
\end{table}

Then we analyzed the 141436 cascades containing more than ten tweets to calculate human-bot interaction. 
We made the following observations:

\begin{itemize}
    \item 1.78\% of all retweets were made from bot Tweets by bots.
    \item 12.04\% of all retweets were made from bot Tweets by humans.
    \item 12\% of all retweets were made from humans Tweets by bots.
    \item 74.18\% of all retweets were made from human Tweets by humans.
\end{itemize}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{1\linewidth}
      \includegraphics[width=\linewidth]{images/bots_furtherretweeted.png}
      \caption{Bots Parents}
      \label{fig:further_retweet_bots}
    \end{subfigure}
    \begin{subfigure}[b]{1\linewidth}
      \includegraphics[width=\linewidth]{images/humans_furtherretweeted.png}
      \caption{Humans Parents}
      \label{fig:further_retweet_humans}
    \end{subfigure}
    \caption{Retweet cascades for tweets made by humans and bots are grouped by their size and visualized. Here, we see 
    that bots dominate cascades started by bots in the early retweets. As the size gets bigger, more humans see it. For 
    human started tweets, their percentage is similar in all sizes}
    \label{fig:further_retweet}
\end{figure}

Next, we compared the cascades started by bots and humans. Figure \ref{fig:further_retweet} can be zoomed 
for detail. In Figure \ref{fig:further_retweet_bots} we can see that the percentage of bots was high for smaller cascades 
started by bots. As the size of the cascade increases, the percentage of bots decreases. 
We made the opposite observation for humans in Figure \ref{fig:further_retweet_humans}. If a tweet receives 
few audiences, they are the followers. Bots are mostly
followed by bots and humans by humans. This finding goes in line with those made by earlier research which
 shows that bots and humans are mostly followed by their type. \par
